---
title: "geograbr"
output:
  md_document:
    variant: markdown_github	
---

# geograbr

A tool for efficiently grabbing usable metadata and experiment data off GEO for use in R.

## Getting experiment metadata and GSE accession numbers

Metadata by experiment accession number can be returned for **all** GEO experiments or **by GPL platform code**

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
comment = ">",
error = FALSE,
tidy = FALSE,
echo = TRUE, 
warning = FALSE, 
message = FALSE,
width=85,
eval = FALSE)
```

```{r load.grabr, eval = TRUE}
devtools::load_all("~/repos/geograbr")
require("XML")
```

```{r retrieve.datasets.01}
## all datasets in GEO
datasets <- geograbr.retrieve.datasets()  ## about 4min
```
```{r retrieve.datasets.02, eval = TRUE}
## all Illumina Infinium 27K DNA methylation datasets
datasets <- geograbr.retrieve.datasets(gpl = c("GPL13534")) ## 5sec
str(datasets)
```


## Download GEO series matix files for by GSE experiment numbers

```{r download.series.files}
## by individual gses
filenames <- geograbr.download.series.files(gses = "GSE94734")

## or for a vector of gses
filenames <- geograbr.download.series.files(gses = datasets$accession[1:6])
```

## Grab sample-level data from downloaded series matrix files

```{r get.samples}

gse94734 <- geograbr.get.samples(filename = "GSE94734_series_matrix.txt.gz")

gses <- lapply(na.omit(filenames), geograbr.get.samples)

```

## Grab experiment data from downloaded series matrix files

Automatically saves data to `.rdata` format at user-defined path

```{r get.data}
test <- geograbr.get.data(filename = "GSE94734_series_matrix.txt.gz")
```
