---
title: "geograbbi: grabbing GEO data efficiently"
author: "Paul Yousefi"
date: "`r Sys.Date()`"
output: 
  rmarkdown::md_document:
    toc: true
    toc_depth: 2
---

# geograbbi

A tool for efficiently grabbing usable metadata and experiment data off GEO for use in R.

## Getting experiment metadata and GSE accession numbers

Metadata by experiment accession number can be returned for **all** GEO experiments or **by GPL platform code**

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
comment = ">",
error = FALSE,
tidy = FALSE,
echo = TRUE, 
warning = FALSE, 
message = FALSE,
width=85,
eval = FALSE)
```

```{r load.grabr, eval = TRUE}
suppressWarnings(suppressPackageStartupMessages(library(geograbbi)))
#require("XML")
```

```{r retrieve.datasets.01}
## all datasets in GEO
datasets <- geograbbi.retrieve.datasets()  ## about 4min
```
```{r retrieve.datasets.02, eval = TRUE}
## all Illumina Infinium 27K DNA methylation datasets
datasets <- geograbbi.retrieve.datasets(gpl = c("GPL13534")) ## 5sec
str(datasets)
```


## Download GEO series matix files for by GSE experiment numbers

```{r download.series.files}
## by individual gses
filenames <- geograbbi.download.series.files(gses = "GSE94734")

## or for a vector of gses
filenames <- geograbbi.download.series.files(gses = datasets$accession[1:6])
```

## Grab sample-level data from downloaded series matrix files

```{r get.samples}

gse94734 <- geograbbi.get.samples(filename = "GSE94734_series_matrix.txt.gz")

gses <- lapply(na.omit(filenames), geograbbi.get.samples)

```

## Grab experiment data from downloaded series matrix files

Automatically saves data to `.rdata` format at user-defined path

```{r get.data}
test <- geograbbi.get.data(filename = "GSE94734_series_matrix.txt.gz")
```
